Selecting Team Members of The Female United States Artistic Gymnastics Team at The 2024 Paris Olympics Based on Performance and Stability To Maximize Success

### 2024-01-15 manuscript/submission.tex
making necessary formatting and figure changes

  Following the 2020 Toyko Olympics where the United States was the only country to have a team medal 
  in all six of the women's artistic gymnastic events, the performance of the United States women's artistic 
  gymnastics team at the 2024 Paris Summer Olympics is once again predicted to be the most watched Olympic 
  dicipline both in the United States and globally. With the whole world anticipating the success of the United 
  States women's team, it is crucial that the team of five women that are selected to represent the United States 
  are the athletes best suited to optimize success and maximize the expected medal count in Paris 2024. As there 
  are no preexisting statistical papers that methodize the selection of athletes for Olympic Gymnastic teams, 
  this paper aims to answer the question of what metric should be used to judge the olympic canidacy of USA 
  gymnasts and to outline a statistical method that can be used presently and in the future to select the ``best'' 
  gymnasts to comprise the United States women's Olympic team in order to maximize success. Using an anatlytical 
  method dependent on mean and standard deviation derived parameters, this paper outlines a procedure for identifying 
  the group of five athletes who will enable the Team USA Olympic Women’s Artistic Gymnastics team to optimize 
  success through numberically evaluating each USA athlete based on both their accuracy and precision in order to 
  more accurately identify the ``best'' team candidates. Through application of this analytical method to present 
  USA female artistic gymnasts, Simone Biles, Shilese Jones, Konnor McClain, Kaliya Lincoln, and Jade Carey can be 
  identified as the the group of five athletes who will enable the Team USA Olympic Women’s Artistic Gymnastics 
  teams to optimize success at the 2024 Paris Summer Olympics.

@@

\citet{JonasRohleder2023} did something...

@@

Currently there is no literature available relevant to predicting gymnastics outcomes at the 2024 
Paris Summer Olympics or at recent olympics. However, there is literature that relates to various aspects of 
the data that will be used to form my predictions. For example, the article ``The Prediction of All-Around 
Event Final Score Based On D and E Score Factors In Women's Artistic Gymnastics'' aims to determine the 
impact of individual apparatus difficulty and execution scores on overall score. 
Expertise from this article could be useful in allowing us to understand how difficulty and execution scores 
play a role in predicting the overall score. Additionally, the paper 
``Performance rating in men’s world elite artistic gymnastics: a status-quo study on scoring 
tendencies at Olympic Games following rule changes'' also can play a similar role in predicting overall 
score. Notably, the paper found that recent Olympic rule changes resulted in significantly enhanced performance 
differentiation with the execution score being the main predictor of the final score. 
Furthermore, in the article ``How Apparatus Difficulty Scores Affect All 
Around Results in Men's Artistic Gymnastics,'' the authors use regression, cluster, and ANOVA analysis to 
analyze how difficulty scores predict overall score differently across different apparatuses. However, in this 
article the authors found that while difficulty scores may have predictive tendancies for some apparatuses, 
the strength of the prediction varied drastically from apparatus to apparatus, with there being little to no 
predictive ability for some apparatuses. Ultimately, the available literature on predicting artistic 
gymnastic scores seemed generally limited to how difficulty and execution scores predict overall score, 
but some aspects of this literature casts doubt on whether difficulty and execution scores should be 
used as predictors at all.

@@

\begin{enumerate}
  \item Calculate and examine the mean and standard deviations of each American female athlete's score 
  on each apparatus.
  \item Calculate the number of observations for each athlete on each apparatus.
  \item Plot the mean versus the standard deviation of the ten athletes with the highest mean scores on each 
  apparatus.
  \begin{enumerate}
    \item Use these plots as a visual to represent the possible scores that could be achielved in each apparatus 
    component of the Olympics.
    \item Using the plots, create a parameter for each apparatus that uses both mean score and standard 
    deviation to identify which athletes are best suited to be considered to compete on behalf of the United 
    States on each apparatus.
    \begin{enumerate}
      \item Select the five athletes with the highest parameter value for each apparatus to identify the ``best'' 
      five athletes that compete on each apparatus based on both their success, a high mean score, and their 
      inconsistency, their standard deviation.
    \end{enumerate}
    \item Identify the athletes with the two largest paramter values as the ``best'' on each apparatus.
  \end{enumerate}
  \item Produce a list of ``target athletes'' for furture computations that contains all athletes that were identified 
  as being in the top five for any apparatus in the last step.
  \begin{enumerate}
    \item Remove any athletes from the ``target athletes'' list that have not competed on all aparatuses.
  \end{enumerate}
  \item Calculate the sum of each ``target athlete's'' mean score for each apparatus and the sum of each athlete's standard 
  deviation for each apparatus.
  \item Plot the mean score sum versus the standard deviation sum for each ``target athlete.''
  \begin{enumerate}
    \item Use this plot as a visual to represent the possible scores that could be achielved in the individual all-around 
    component of the Olympics.
  \end{enumerate}
  \item Calculate the summed mean score and standard deviation sum of every combination of three athletes using the previously 
  calculated sum of each ``target athlete's'' mean score for each apparatus and the sum of each athlete's standard 
  deviation for each apparatus.
  \item Plot the summed mean score versus the standard deviation for the ten combinations of three athletes with the 
  highest summed mean score.
  \begin{enumerate}
    \item Use this plot as a visual to represent the possible scores that could be achielved in the all-around team finals 
   component of the Olympics.
  \end{enumerate}
  \item Calculate the summed mean score and standard deviation sum of every combination of four athletes using the previously 
  calculated sum of each ``target athlete's'' mean score for each apparatus and the sum of each athlete's standard 
  deviation for each apparatus.
  \item Plot the summed mean score versus the standard deviation for the ten combinations of four athletes with the 
  highest summed mean score.
  \begin{enumerate}
    \item Use this plot as a visual to represent the possible scores that could be achielved in the all-around team qualifying  
   round of the Olympics.
  \end{enumerate}
  \item Visully exammine each produced plot to build a final team of five in an effort to include in 
  the final team the best three candidates for the team all-around, the best candidate for the fourth qualifying position, 
  the best two candidates for the individual all-around, and the two athletes that are superior on each apparatus.

\end{enumerate}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.7]{FemaleAthleteApparatus.pdf}
  \caption{Top 10 Female Candidates for the Olympic Balance Beam, Floor Exercise, Uneven Bars, and Vault Apparatus}
  \label{fig:FA}
\end{figure}

\begin{table}
  \caption{Top ten female candidates for each apparatus based on mean score alone and 
  their $\mu - \alpha \sigma$ parameter values}
  \label{tab:tableBBP}
\centering
\begin{tabular}[t]{llllllll}
 \toprule
  \multicolumn{2}{c}{BB} & \multicolumn{2}{c}{FX} & \multicolumn{2}{c}{UB} & \multicolumn{2}{c}{VT}\\
  \cmidrule(lr){1-2}\cmidrule(lr){3-4}\cmidrule(lr){5-6}\cmidrule(lr){7-8}
Athlete & $\mu - \sigma$ & Athlete & $\mu - \sigma$ & Athlete & $\mu - 0.5 \sigma$ & Athlete & $\mu - 0.5 \sigma$\\
\midrule
biles & 14.38 & biles & 14.55 & jones & 14.26 & Biles & 14.78\\
\midrule
mcclain & 14.14 & lincoln & 13.73 & biles & 14.17 & Carey & 14.36\\
\midrule
lee & 13.69 & carey & 13.41 & wong & 13.59 & Mcclain & 14.20\\
\midrule
sumanasekera & 12.95 & dicello & 13.4 & matthews & 13.50 & Jones & 14.20\\
\midrule
blakely & 12.78 & chiles & 13.38 & blakely & 13.49 & Chiles & 14.11\\
\midrule
jones & 12.58 & jones & 13.29 & greaves & 13.47 & sumanasekera & 14.07\\
\midrule
roberson & 12.50 & mcclain & 13.21 & miller & 13.38 & blakely & 14.05\\
\midrule
wong & 12.48 & wong & 13.16 & chiles & 13.34 & dicello & 13.88\\
\midrule
dicello & 12.45 & roberson & 12.98 & zhou & 13.28 & richardson & NA\\
\midrule
frazier & NA & frazier & NA & walker & NA & torry & NA\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.55]{FemaleMultipleAllAround.pdf}
  \caption{The best candidates for the female individual all-around and team all around medals}
  \label{fig:FAA}
\end{figure}

@@

\section{Appendix: Application of the Mean and Standard Deviation Derived Parameters Methods For The Male 
United States Artistic Gymnastics Team at The 2024 Paris Olympics}

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{MaleAthleteApparatus.pdf}
  \caption{Top 10 Male Candidates for the Olympic Floor Exercise, Pommel Horse, Still Rings, Vault, Parallel Bars, 
  and High Bar Apparatus}
  \label{fig:MA}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=0.7]{MaleMultipleAllAround.pdf}
  \caption{The best candidates for the male individual all-around and team all around medals}
  \label{fig:MAA}
\end{figure}

>>>>

  With the whole world anticipating the success of the United States women's artistic gymnastics team, 
  it is crucial that the team of five women that are selected to represent the United States 
  are the athletes best suited to optimize success and maximize the expected medal count in Paris 2024. As there 
  are no preexisting statistical papers that methodize the selection of athletes for Olympic Gymnastic teams, 
  this paper aims to answer the question of what metric should be used to judge the Olympic canidacy of USA 
  gymnasts and to outline a statistical method that can be used presently and in the future to select the ``best'' 
  gymnasts to comprise the United States women's Olympic team in order to maximize success. Using an anatlytical 
  method dependent on mean and standard deviation derived parameters, this paper outlines a procedure for identifying 
  the group of five athletes who will enable the Team USA Olympic Women’s Artistic Gymnastics team to optimize 
  success through numberically evaluating each USA athlete based on both their performance and stability in order to 
  more accurately identify the ``best'' team candidates. Through application of this analytical method to present 
  USA female artistic gymnasts, Simone Biles, Shilese Jones, Konnor McClain, Kaliya Lincoln, and Jade Carey can be 
  identified as the the group of five athletes who will enable the Team USA Olympic Women’s Artistic Gymnastics 
  teams to optimize success at the 2024 Paris Summer Olympics.

@@

Following the 2020 Toyko Olympics where the United States was the only country to have a team medal 
in all six of the women's artistic gymnastic events, the performance of the United States women's artistic 
gymnastics team at the 2024 Paris Summer Olympics is once again predicted to be the most watched Olympic 
dicipline both in the United States and globally. With the whole world anticipating the success of the United 
States women's team, it is crucial that the team of five women that are selected to represent the United States 
are the athletes best suited to optimize success and maximize the expected medal count in Paris 2024.

@@

Currently there is no literature available relevant to predicting gymnastics outcomes at the 2024 
Paris Summer Olympics or at recent olympics. However, there is literature that relates to various aspects of 
the data that will be used to form my predictions. For example, the article \citet{ScienceOfGymnastics2021} aims to determine the 
impact of individual apparatus difficulty and execution scores on overall score. 
Expertise from this article could be useful in allowing us to understand how difficulty and execution scores 
play a role in predicting the overall score. Additionally, the article \citet{JonasRohleder2023} 
also can play a similar role in predicting overall 
score. Notably, the paper found that recent Olympic rule changes resulted in significantly enhanced performance 
differentiation with the execution score being the main predictor of the final score. 
Furthermore, in the article \citet{ApparatusDifficulty2010} the authors use regression, cluster, and ANOVA analysis to 
analyze how difficulty scores predict overall score differently across different apparatuses. However, in this 
article the authors found that while difficulty scores may have predictive tendancies for some apparatuses, 
the strength of the prediction varied drastically from apparatus to apparatus, with there being little to no 
predictive ability for some apparatuses. Ultimately, the available literature on predicting artistic 
gymnastic scores seemed generally limited to how difficulty and execution scores predict overall score, 
but some aspects of this literature casts doubt on whether difficulty and execution scores should be 
used as predictors at all.

@@

\begin{enumerate}
  \item Determine the best individual apparatus candidates.
  \begin{enumerate}
    \item Calculate and examine the mean and standard deviations of each American female athlete's score 
    on each apparatus.
    \item Calculate the number of observations for each athlete on each apparatus.
    \item Plot the mean versus the standard deviation of the ten athletes with the highest mean scores on each 
    apparatus.
    \begin{enumerate}
      \item Use these plots as a visual to represent the possible scores that could be achielved in each apparatus 
     component of the Olympics.
     \item Using the plots, create a parameter for each apparatus that uses both mean score and standard 
     deviation to identify which athletes are best suited to be considered to compete on behalf of the United 
     States on each apparatus.
     \begin{enumerate}
       \item Select the five athletes with the highest parameter value for each apparatus to identify the ``best'' 
        five athletes that compete on each apparatus based on both their success, a high mean score, and their 
        inconsistency, their standard deviation.
     \end{enumerate}
     \item Identify the athletes with the two largest paramter values as the ``best'' on each apparatus.
    \end{enumerate}
  \end{enumerate}
  \item Determine the best individual all-around candidates.
  \begin enumerate
   \item Produce a list of ``target athletes'' for furture computations that contains all athletes that were identified 
   as being in the top five for any apparatus in the last step.
   \begin{enumerate}
     \item Remove any athletes from the ``target athletes'' list that have not competed on all aparatuses.
   \end{enumerate}
   \item Calculate the sum of each ``target athlete's'' mean score for each apparatus and the sum of each athlete's standard 
   deviation for each apparatus.
    \item Plot the mean score sum versus the standard deviation sum for each ``target athlete.''
   \begin{enumerate}
     \item Use this plot as a visual to represent the possible scores that could be achielved in the individual all-around 
     component of the Olympics.
   \end{enumerate}
  \end{enumerate}
  \item Determine the best team all-around candidates.
  \begin{enumerate}
   \item Calculate the summed mean score and standard deviation sum of every combination of three athletes using the previously 
   calculated sum of each ``target athlete's'' mean score for each apparatus and the sum of each athlete's standard 
   deviation for each apparatus.
   \item Plot the summed mean score versus the standard deviation for the ten combinations of three athletes with the 
   highest summed mean score.
   \begin{enumerate}
     \item Use this plot as a visual to represent the possible scores that could be achielved in the all-around team finals 
    component of the Olympics.
   \end{enumerate}
   \item Calculate the summed mean score and standard deviation sum of every combination of four athletes using the previously 
   calculated sum of each ``target athlete's'' mean score for each apparatus and the sum of each athlete's standard 
   deviation for each apparatus.
   \item Plot the summed mean score versus the standard deviation for the ten combinations of four athletes with the 
   highest summed mean score.
   \begin{enumerate}
      \item Use this plot as a visual to represent the possible scores that could be achielved in the all-around team qualifying  
     round of the Olympics.
   \end{enumerate}
  \end{enumerate}
\item Visully exammine each produced plot to build a final team of five in an effort to include in 
    the final team the best three candidates for the team all-around, the best candidate for the fourth qualifying position, 
    the best two candidates for the individual all-around, and the two athletes that are superior on each apparatus.
\end{enumerate}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.7]{FinalFemaleApparatusPlot.pdf}
  \caption{Top 10 Female Candidates for the Olympic Balance Beam, Floor Exercise, Uneven Bars, and Vault Apparatus.}
  \label{fig:FA}
\end{figure}

\begin{table}
  \caption{Top ten female candidates for each apparatus based on mean score alone and 
  their $\mu - \alpha \sigma$ parameter values.}
  \label{tab:tableBBP}
\centering
\begin{tabular}[t]{llllllll}
 \toprule
  \multicolumn{2}{c}{BB} & \multicolumn{2}{c}{FX} & \multicolumn{2}{c}{UB} & \multicolumn{2}{c}{VT}\\
  \cmidrule(lr){1-2}\cmidrule(lr){3-4}\cmidrule(lr){5-6}\cmidrule(lr){7-8}
Athlete & $\mu - \sigma$ & Athlete & $\mu - \sigma$ & Athlete & $\mu - 0.5 \sigma$ & Athlete & $\mu - 0.5 \sigma$\\
\midrule
biles & 14.38 & biles & 14.55 & jones & 14.26 & Biles & 14.78\\
\midrule
mcclain & 14.14 & lincoln & 13.73 & biles & 14.17 & Carey & 14.36\\
\midrule
lee & 13.69 & carey & 13.41 & wong & 13.59 & Mcclain & 14.20\\
\midrule
sumanasekera & 12.95 & dicello & 13.4 & matthews & 13.50 & Jones & 14.20\\
\midrule
blakely & 12.78 & chiles & 13.38 & blakely & 13.49 & Chiles & 14.11\\
\midrule
jones & 12.58 & jones & 13.29 & greaves & 13.47 & sumanasekera & 14.07\\
\midrule
roberson & 12.50 & mcclain & 13.21 & miller & 13.38 & blakely & 14.05\\
\midrule
wong & 12.48 & wong & 13.16 & chiles & 13.34 & dicello & 13.88\\
\midrule
dicello & 12.45 & roberson & 12.98 & zhou & 13.28 & richardson & NA\\
\midrule
frazier & NA & frazier & NA & walker & NA & torry & NA\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.55]{FinalFemaleAllAroundPlot.pdf}
  \caption{The best candidates for the female individual all-around and team all around medals.}
  \label{fig:FAA}
\end{figure}

@@

\section{Application of Methods To The Male USA Artistic Gymnastics Team}

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{FinalMaleApparatusPlot.pdf}
  \caption{Top 10 Male Candidates for the Olympic Floor Exercise, Pommel Horse, Still Rings, Vault, Parallel Bars, 
  and High Bar Apparatus.}
  \label{fig:MA}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=0.7]{FinalMaleAllAroundPlot.pdf}
  \caption{The best candidates for the male individual all-around and team all around medals.}
  \label{fig:MAA}
\end{figure}
